A Typed Programmatic Interface to Contracts on the Blockchain

### 2021-02-21 papers/execution-model/model.tex
intro & conclusion

{[Ultra-brief intro to blockchain]}

Smart contracts rarely live in isolation. To reach their full
potential, it should be possible to integrate them with traditional
software systems and to enable input from outside the blockchain. Such
input requires trusted intermediaries that run so-called oracles \cite{oracle-patterns,call-action-oracle}. A
smart contract registers a request and a callback with an oracle. The
oracle invokes the callback as soon as an answer is available.

We present two examples where integration of smart contracts is needed
and propose a typed API for making such integration type safe.

@@

>>>>

First generation blockchains were primarily geared towards supporting
cryptocurrencies. The most prominent system of this kind is Bitcoin
\cite{bitcoin-whitepaper}. Though Bitcoin already features a
rudimentary programming language called Script, second generation
blockchains like the Ethereum blockchain \cite{eth-whitepaper} feature
Turing-complete programming facilities, called 
Smart Contracts. They provide robust trustworthy distributed computing
facilities even though the programs run on a peer-to-peer network with
untrusted participants. Each peer in the network runs the same program
and uses cryptographic methods to check the results among the other
peers and to create a persistent ledger of all transactions, the
blockchain, thus ensuring the integrity of the results. Third
generation blockchains, like Tezos \cite{tezos-whitepaper}, are even adaptable to
new requirements without breaking participating peers (no ``hard
forks'' required).

The strength of programs on the blockchain is also their
weakness. They are fully deterministic in that they can only depend on
data that is stored on the chain and on contract invocations, which are
also stored on the chain. Moreover, the entire data as well as all
transactions are public. These properties make it hard to react to
external stimuli like time triggers or events like an exchange rate or
a share price hitting a limit.

As smart contracts rarely live in isolation, it should be possible to
integrate them with traditional software systems and to enable
triggering contracts from outside the blockchain. Such triggers
require trusted intermediaries that provide so-called oracles \cite{oracle-patterns,call-action-oracle}. A
smart contract registers a request and a callback with an oracle. The
oracle invokes the callback as soon as an answer is available.

We present a typed API for invoking contracts from OCaml
programs. Such an API with a reliable computing facility and network
connection is sufficient to self provision an oracle and to avoid
having to trust an oracle provider. Moreover, our approach provides a
type-safe facility rather than oracles which are intrisically untyped
because they aim to provide a generic service.

@@

\section{Conclusion}
\label{sec:conclusion}

We present a first step towards a typed API for smart contracts on the
Tezos blockchain. Our formalization enables us to establish basic
properties of the interaction between ordinary programs and smart
contracts. We see ample scope for future work to provide a
higher-level interface that exploits the similarities between
blockchain programming and concurrent programs.
